---
title: "Homework 6"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
library("leaps")

knitr::opts_chunk$set(
  fig.width = 9,
  fig.asp = .8,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
set.seed(1)
```

## Problem 1

### Import and manipulate data

```{r read_data, message = FALSE, warning = FALSE}
homicide_df = 
  read_csv("./homicide-data.csv") %>% 
  mutate(city_state = str_c(city, ", ", state),
         resolved = as.numeric(disposition == "Closed by arrest")) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", 
                             "Kansas City, MO", "Tulsa, AL"))) %>% 
  mutate(victim_race = ifelse(victim_race != "White", "Non-White", "White"),
         victim_race = fct_relevel(victim_race, "White"),
         victim_age = as.numeric(victim_age)) %>% 
  select(city_state, resolved, everything())

homicide_df
str(homicide_df)
```

This dataset contains `r nrow(homicide_df)` observations, where each observation is a case of homicide. There are 14 variables in this dataset, including the location of the killing, whether the homicide was solved and basic demographic information about each victim.

### For the city of Baltimore, MD

First, fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 
```{r logistic_regression}
baltimore_df = filter(homicide_df, city == "Baltimore")

fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy()
fit_logistic
```

Homicides in which the victim is non-white are substantially less likely to be resolved than those in which the victim is white. Homicides in which the victim is male are significantly less likely to be resolved than those in which the victim is female. The effect of age is statistically significant, but careful data inspections should be conducted before interpreting too deeply.

Then, obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed:

```{r odds_ratio}
fit_logistic %>% 
  filter(term == "victim_raceNon-White") %>% 
  mutate(OR = exp(estimate), 
         lower_limit = exp(estimate - qnorm(0.975)*std.error), 
         upper_limit = exp(estimate + qnorm(0.975)*std.error)) %>% 
  select(term, OR, lower_limit, upper_limit) %>% 
  knitr::kable(digits = 3)
```

The estimate of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed is 0.441 and the 95% confidence interval is (0.313, 0.620).

### For each city

Create a dataframe with estimated ORs and CIs for each city:

```{r ors_cis}
or_df = 
  homicide_df %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(resolved_logistic = 
         map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),
         resolved_logistic = map(resolved_logistic, broom::tidy)) %>% 
  select(city_state, resolved_logistic) %>% 
  unnest() %>% 
  filter(term == "victim_raceNon-White") %>% 
  mutate(OR = exp(estimate), 
         lower_limit = exp(estimate - qnorm(0.975)*std.error), 
         upper_limit = exp(estimate + qnorm(0.975)*std.error)) %>% 
  select(city_state, OR, lower_limit, upper_limit) 
or_df
```

### Odds ratio plot

```{r or_plot}
or_df %>% 
  mutate(city_state = fct_reorder(city_state, desc(OR))) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_limit, ymax = upper_limit)) + 
  geom_hline(yintercept = 1, alpha = 0.5, color = "red") +
  coord_flip() +
  labs(title = "Estimated ORs and 95% CIs for solving homicides", 
       y = "Odds ratio for solving homicides", 
       x = "City and State", 
       caption = "Data from the Washington Post") 
```

The above plot shows the estimated ORs and CIs for solving homicides comparing non-white victims to white victims for each city, ordered from most to least. According to the plot, the estimated adjusted OR in most cities for solving homicides comparing non-white to white victims is less than 1, which means in most cities, homicides in which the victim is non-white are less likely to be resolved than those in which the victim is white. Tampa, Florida and Birmingham, Alabama are the only two cities with estimated ORs greater than one, however, both estimates have very wide confidence intervals. In addition, about half of the cities have confidence intervals that include 1.

## Problem 2

### Import data

Load and clean the data for regression analysis:

```{r clean_data}
bwt_data = read_csv("./birthweight.csv")
bwt_data =
  bwt_data %>% 
  mutate(babysex = as.factor(recode(babysex, '1' = 'Male', '2' = 'Female')), 
         frace = as.factor(recode(frace, `1` = 'White', `2` = 'Black', 
                               `3` = 'Asian', `4` = 'Puerto Rican', 
                               `8` = 'Other', `9` = 'Unknown')), 
         malform = as.factor(ifelse(malform == 1, "present", "absent")), 
         mrace = as.factor(recode(mrace, `1` = 'White', `2` = 'Black',
                               `3` = 'Asian', `4` = 'Puerto Rican',
                               `8` = 'Other'))) %>% 
  select(bwt, everything())
bwt_data
```

Check for missing data:
```{r check_missing}
sum(is.na(bwt_data))
```

This dataset contains `r nrow(bwt_data)` observations with no missing data. I converted numeric to factor where appropriate.

### Model building

Exploratory analyses:

```{r}
bwt_data %>% 
  ggplot(aes(x = bwt)) + 
  geom_histogram() + 
  labs(title = "Histogram plot of birthweight",
       x = "Birthweight", 
       y = "Count")
```

From the histogram plot, baby's birth weight appears to be normally distributed which satisfies the linear regression model assumptions.

Propose a regression model for birthweight:

```{r modeling}
mult.fit = lm(bwt ~ ., data = bwt_data)
step(mult.fit, direction = 'backward')

model_my = lm(bwt ~ babysex + bhead + blength + delwt + fincome + 
              gaweeks + mheight + mrace + parity + ppwt + smoken, data = bwt_data)
summary(model_my)
```

*Description of my modeling process:* First, I checked the distribution of the outcome -- birth weight and it turned out to be normally distributed, so I decided to build a linear regression model. Then I used stepwise regression procedure to select variables and a single regression model is identified as the 'best' model in the end of the stepwise regression. 'Step' function uses AIC criterion for variable selection and the default option is 'backward'. I picked the final model as my model, which includes 11 predictors -- baby's sex, baby's head circumference at birth (centimeters), baby's length at birth (centimeteres), mother's weight at delivery (pounds), family monthly income (in hundreds, rounded), gestational age in weeks, mother's height (inches), mother's race, number of live births prior to this pregnancy, mother's pre-pregnancy weight (pounds), and average number of cigarettes smoked per day during pregnancy.

Show a plot of model residuals against fitted values:

```{r residuals_plot}
bwt_data %>% 
  add_predictions(model_my) %>% 
  add_residuals(model_my) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
        title = "Plot of Model Residuals VS Fitted Values",
        x = "Predictions",
        y = "Residuals"
      ) 
```

The above scatter plot shows the relationship between predicted values and residuals for the my model. It appears that the equality of variance assumption is not holding when the predicted values of birth weight are very low and that outliers are present. However, for most observatons, residuals form a horizontal (linear) band around zero, which indicates equal variance.